Almost Orthogonal vectors

### 2014-11-26 project3/paper/paper.tex
Removed unnecessary figure
More changes to paper

% Graphics
%\usepackage[dvips]{graphicx,epsfig,color}
%\usepackage{subfigure}
%\usepackage{pst-all}
%\usepackage{pstricks-add}
\usepackage{hyperref}  % can only be used with pdflatex - gives hyperlinks
\usepackage{bm} % bold math font
\usepackage{bbm}

@@

\begin{figure}
    \centering
    \begin{tikzpicture}[yscale=-1] 
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % origin label
        \node at (-0.5, 0.3) {\text{$(0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
        % circle
        \draw (0,0) circle (3cm);
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (1.5, -2.59)[blue,fill=blue] circle (0.1cm);
        \draw (-1.5, -2.59)[blue,fill=blue] circle (0.1cm);

@@

Our arguments will hinge on the fact that for all $\theta \in [0, \pi/2]$, the function $\cos \theta$ is decreasing. Without loss of generality, let us assume that $\theta_1 \geq \theta_2 \geq \ldots \theta_n$. Hence we want to prove that $\epsilon$ is equal to $\cos^2 \theta_n$.

We split our proof into two cases,
\begin{itemize}
\item $\theta_1, \theta_2, \ldots, \theta_n \in [0, \pi/2]$: This is easy to see from the fact that $\cos \theta$ is a decreasing function in $\theta$ if $\theta \in [0, \pi/2]$.

\item One of $\theta_1, \theta_2, \dots, \theta_n$ is greater than $\pi/2$:
\begin{figure}
    \centering
    \begin{tikzpicture}[yscale=-1] 
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % origin label
        \node at (-0.5, 0.3) {\text{$(0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
        % circle
        \draw (0,0) circle (3cm);
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (2.42, -1.76)[blue,fill=blue] circle (0.1cm);
        \draw (0.93, -2.85)[blue,fill=blue] circle (0.1cm);
        \draw (-2.42, -1.76)[blue,fill=blue] circle (0.1cm);
        \draw (-0.93, -2.85)[blue,fill=blue] circle (0.1cm);

        \node at (4, -0.3) {\text{$x_1 = [1,0]$}};
        \node at (4.5, -1.76) {\text{$x_2 = [\cos \pi/5,\sin \pi/5]$}};
        \node at (3.43, -3) {\text{$x_3 = [\cos 2\pi/5,\sin 2\pi/5]$}};
         \node at (-4.5, -1.76) {\text{$x_4 = [\cos 4\pi/5,\sin 4\pi/5]$}};
        \node at (-3.43, -3) {\text{$x_5 = [\cos 3\pi/5,\sin 3\pi/5]$}};
    \end{tikzpicture}
    \caption{A configuration of five unit vectors that produce the optimum $\epsilon$ for $n=2, N=5$}
\end{figure}

@@

It is easy to see that for all $i \in \{2, 3, \ldots, n-1\}$, $\cos^2 \theta_i < \cos^2 \theta_n$, from which we can conclude that in this case too, $\epsilon$ is equal to $\cos^2 \theta_n$.

@@

\begin{proof}
The proof is easy to see given the above lemma.

Again without loss of generality, we can assume that the vectors $x_1, x_2, \ldots, x_N$ are on or above the $x$-axis, and that $x_1 = [1,0]^T$ and that $x_{N+1} = [-1,0]^T$.

@@

Figure $2$ shows an optimum configuration for $n=2$ and $N=5$.
\begin{figure}
    \centering
    \begin{tikzpicture}[yscale=-1] 
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % origin label
        \node at (-0.5, 0.3) {\text{$(0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
        % circle
        \draw (0,0) circle (3cm);
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (-1.5, 2.59)[blue,fill=blue] circle (0.1cm);
        \draw (-1.5, -2.59)[blue,fill=blue] circle (0.1cm);

        % \draw [thick,-,blue, dotted] (-3, 0) -- (0, 0);
        % \draw [thick,-,blue, dotted] (1.5, -2.59) -- (0, 0);
        % \draw [thick,-,blue, dotted] (1.5, 2.59) -- (0, 0);
        \draw [thick,-,blue] (3, 0) -- (0, 0);
        \draw [thick,-,blue] (-1.5, 2.59) -- (0, 0);
        \draw [thick,-,blue] (-1.5, -2.59) -- (0, 0);
    \end{tikzpicture}
    \caption{A configuration of three unit vectors that produce the optimum $\epsilon$ for $n=2, N=3$}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \begin{tikzpicture}[yscale=-1] 
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % origin label
        \node at (-0.5, 0.3) {\text{$(0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
        % circle
        \draw (0,0) circle (3cm);
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (-1.5, 2.59)[blue,fill=blue] circle (0.1cm);
        \draw (-1.5, -2.59)[blue,fill=blue] circle (0.1cm);

>>>>

% Graphics
%\usepackage[dvips]{graphicx,epsfig,color}
%\usepackage{subfigure}
%\usepackage{pst-all}
%\usepackage{pstricks-add}
\usepackage{hyperref}  % can only be used with pdflatex - gives hyperlinks
\usepackage{bm} % bold math font
\usepackage{bbm}
\usepackage{pgfplots}

@@

\begin{figure}[!h]
    \centering
    \begin{tikzpicture}[yscale=-1] 
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % origin label
        \node at (-0.5, 0.3) {\text{$(0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
        % circle
        \draw (0,0) circle (3cm);
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (1.5, -2.59)[blue,fill=blue] circle (0.1cm);
        \draw (-1.5, -2.59)[blue,fill=blue] circle (0.1cm);

@@

\begin{figure}[!h]
	\centering
	\begin{tikzpicture}
		\begin{axis}[%
			axis x line=center, axis y line=center,
			width=10cm,
			height=4cm,
			scale only axis,
			xmin=-5,
			xmax=5,
			xtick={1.57,  3.14, 4.71},
			xticklabels={, $\pi$, },
			extra x ticks={-4.71, -3.14, -1.57},
			extra x tick labels={, $-\pi$,},
			extra x tick style={
			    xticklabel style={yshift=0.5ex, anchor=south}
			},
			ymin=-1.4,
			ymax=1.4,
			ytick={-1,  0,  1}]]
			\addplot[domain=-2*pi:2*pi,smooth] (\x,{cos(\x r)});
		\end{axis}
	\end{tikzpicture}
	\caption{Plot of $\cos \theta$ versus $\theta$}
\end{figure}

Our arguments will hinge on the fact that for all $\theta \in [0, \pi/2]$, the function $\cos \theta$ is decreasing -- this is easy to see from Figure $2$. Without loss of generality, let us assume that $\theta_1 \geq \theta_2 \geq \ldots \geq\theta_n$. Hence we want to prove that $\epsilon$ is equal to $\cos^2 \theta_n$.

We split our proof into two cases,
\begin{itemize}
\item $\theta_1, \theta_2, \ldots, \theta_n \in [0, \pi/2]$: In this case, it is easy to see that $\epsilon = \cos^2 \theta_n$ from the fact that $\cos \theta$ is a decreasing function in $\theta$ if $\theta \in [0, \pi/2]$.

\item One of $\theta_1, \theta_2, \dots, \theta_n$ is greater than $\pi/2$:

@@

In addition, we see that for all $i \in \{2, 3, \ldots, n-1\}$, $\cos^2 \theta_i < \cos^2 \theta_n$, from which we can conclude that in this case too, $\epsilon$ is equal to $\cos^2 \theta_n$.

@@

\begin{proof}
The proof of this theorem is similar to the proof for the specific case of $N=3$.

Again without loss of generality, we can assume that the vectors $x_1, x_2, \ldots, x_N$ are on or above the $x$-axis, and that $x_1 = [1,0]^T$ and that $x_{N+1} = [-1,0]^T$ -- if any vector $x_i$ were not above the $x$-axis, then we could just consider its reverse.

@@

\begin{figure}[!h]
    \centering
    \begin{tikzpicture}[yscale=-1] 
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % origin label
        \node at (-0.5, 0.3) {\text{$(0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
        % circle
        \draw (0,0) circle (3cm);
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (2.42, -1.76)[blue,fill=blue] circle (0.1cm);
        \draw (0.93, -2.85)[blue,fill=blue] circle (0.1cm);
        \draw (-2.42, -1.76)[blue,fill=blue] circle (0.1cm);
        \draw (-0.93, -2.85)[blue,fill=blue] circle (0.1cm);

        \node at (4, -0.3) {\text{$x_1 = [1,0]$}};
        \node at (4.5, -1.76) {\text{$x_2 = [\cos \pi/5,\sin \pi/5]$}};
        \node at (3.43, -3) {\text{$x_3 = [\cos 2\pi/5,\sin 2\pi/5]$}};
         \node at (-4.5, -1.76) {\text{$x_5 = [\cos 4\pi/5,\sin 4\pi/5]$}};
        \node at (-3.43, -3) {\text{$x_4 = [\cos 3\pi/5,\sin 3\pi/5]$}};
    \end{tikzpicture}
    \caption{A configuration of five unit vectors that produce the optimum $\epsilon$ for $n=2, N=5$}
\end{figure}

\begin{figure}[!h]
    \centering
    \begin{tikzpicture}[yscale=-1] 
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % origin label
        \node at (-0.5, 0.3) {\text{$(0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
        % circle
        \draw (0,0) circle (3cm);
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (-2.42, 1.76)[blue,fill=blue] circle (0.1cm);
        \draw (0.93, -2.85)[blue,fill=blue] circle (0.1cm);
        \draw (-2.42, -1.76)[blue,fill=blue] circle (0.1cm);
        \draw (0.93, 2.85)[blue,fill=blue] circle (0.1cm);

        \draw [thick,-,blue] (3, 0) -- (0, 0);
        \draw [thick,-,blue] (-2.42, 1.76) -- (0, 0);
        \draw [thick,-,blue] (0.93, -2.85) -- (0, 0);
        \draw [thick,-,blue] (-2.42, -1.76) -- (0, 0);
        \draw [thick,-,blue] (0.93, 2.85) -- (0, 0);

        \node at (4, -0.3) {\text{$x_1 = [1,0]$}};
        \node at (-4.5, 1.76) {\text{$x'_2 = [\cos \pi/5,\sin \pi/5]$}};
        \node at (3.43, -3) {\text{$x_3 = [\cos 2\pi/5,\sin 2\pi/5]$}};
         \node at (-4.5, -1.76) {\text{$x_5 = [\cos 4\pi/5,\sin 4\pi/5]$}};
        \node at (3.43, 3) {\text{$x'_4 = [\cos 3\pi/5,\sin 3\pi/5]$}};

        \draw [thick,-,red,dashed] (3, 0) -- (0.93, 2.85);
        \draw [thick,-,red,dashed] (3, 0) -- (0.93, -2.85);
        \draw [thick,-,red,dashed] (0.93, 2.85) -- (-2.42, 1.76);
        \draw [thick,-,red,dashed] (0.93, -2.85) -- (-2.42, -1.76);
        \draw [thick,-,red,dashed] (-2.42, 1.76) -- (-2.42, -1.76);
    \end{tikzpicture}
    \caption{A configuration of five unit vectors that produce the optimum $\epsilon$ for $n=2, N=5$ with $x_2$ and $x_4$ from the previous figure reversed}
\end{figure}

Figure $3$ shows an optimum configuration for $n=2$ and $N=5$. Note by taking reflections of $x_2$ and $x_4$ instead, we get five vectors that form a regular $5$-gon -- this can be seen in Figure $4$. This can be done for all odd $N$; unfortunately, even $N$ do not produce such nice polygons.

@@

\begin{figure}[!h]
    \centering
    \begin{tikzpicture}[yscale=-1] 
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % origin label
        \node at (-0.5, 0.3) {\text{$(0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
        % circle
        \draw (0,0) circle (3cm);
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (-1.5, 2.59)[blue,fill=blue] circle (0.1cm);
        \draw (-1.5, -2.59)[blue,fill=blue] circle (0.1cm);
